---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '# '
---

# 🚧 [Part 3: Networking]{style="font-size: 1.05em; font-weight: bold;"} {#sec-part-two}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience."
)
```

This section provides and introduction to Linux networking. We'll cover some foundational theory and hands-on configuration on Ubuntu (22.04+) and RHEL 9 operating systems. Each chapter includes a **Common Commands** box highlighting distro-specific tools and syntax.

---

##  Networking Fundamentals

* **OSI & TCP/IP Models**
* **IPv4 vs. IPv6 Addressing**
* **MAC Addresses & ARP**
* **Common Protocols (ICMP, DHCP, DNS, HTTP/S)**

---

### Network Configurations

* **DHCP vs. Static Assignment**    
* **Ubuntu: Netplan**     
    - `/etc/netplan/*.yaml` syntax    
    - `sudo netplan apply`

* **RHEL 9: NetworkManager & legacy scripts**   
    - `nmcli` CLI: `nmcli connection add/modify`    
    - `nmtui` TUI   
    - (Optional) `/etc/sysconfig/network-scripts/ifcfg-*`   

* **Bringing Interfaces Up/Down**


---

### Domain Name System 

* **DNS & Name Resolution**   
* **`/etc/hosts` Basics**   
* **Resolver Configuration**    
    - Ubuntu: `systemd-resolved`, `/etc/resolv.conf` symlink    
    - RHEL 9: direct `/etc/resolv.conf`   

* **Testing DNS**
* **Caching & Stub Resolvers**


---

### Routing & IP Management

* **Static Routes**
* **Policy-Based Routing (ip rule)**
* **Enabling IP Forwarding**
* **VRFs & Network Namespaces**


---

## Firewalling & Network Security

* **Ubuntu UFW (Uncomplicated Firewall)**   
    - `sudo ufw enable/disable`   
    - Profile management    

* **RHEL 9 firewalld**
    - `sudo firewall-cmd --zone=public --add-port=80/tcp`   
    - Permanent vs. runtime rules   

* **iptables vs. nftables**   
* **SELinux & Network Access**    

---

### Monitoring & Troubleshooting

* **Connectivity Tests (ping, traceroute, mtr)**
* **Socket & Port Inspection (ss, netstat)**
* **Packet Capture (tcpdump, tshark)**
* **Interface Statistics (ethtool, iftop, nload)**
* **Logs & Systemd Journal**

---

### VPNs, Encryption, & Tunnels 

* **SSH Tunnels & SOCKS Proxies**
* **OpenVPN Setup**
* **WireGuard Quick-Start**
* **IPsec & strongSwan**

---

### Linux Virtual & Container Networking

* **Bridge & VLAN Interfaces**
* **Docker & Podman Networks**
* **CNI Plugins & Kubernetes Basics**
* **MACVLAN, NAT & Host Networking Modes**

