---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
execute: 
  eval: false
---

# Permissions {#sec-perms}

```{r}
#| label: setup
#| include: false
#| eval: true
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Every file and folder on Linux has rules about who can see it, change it, or run it. Those rules are called **permissions**, and they’re what keep data safe and organized. In this chapter, we’ll cover:

1.  What permissions are and why they're important\
2.  How to read and understand permission settings\
3.  How to change permissions with simple commands\
4.  A few handy examples and troubleshooting tips

## Why do we need permissions?

**Security and privacy**

Imagine you share a computer with someone else. You don’t want them accidentally deleting your photos, and they don’t want you messing with their work documents. Permissions keep each person’s files separate and safe.

**Accident prevention**

Even on your own machine, permissions protect important system files. If you could delete or change everything without restriction, you might break your system by mistake.

**Controlled sharing**

When you do want to share, permissions let you decide exactly who can view or edit a file or folder.

### Permission Categories

Linux assigns permissions in three categories for every file and folder:

1.  **Owner** (the user who created or “owns” the file)
2.  **Group** (a set of users who share access rights)
3.  **Others** (everyone else on the system)

There are also three types of actions:

-   **Read (`r`)**: View the contents of a file or list a folder’s contents
-   **Write (`w`)**: Change a file or add/remove files in a folder
-   **Execute (`x`)**: Run a file as a program or “enter” a folder to work inside

### View Permissions

Open a terminal and type:

```{bash}
ls -l
```

You’ll see output like this:

```{bash}
# total 8
# drwxr-xr-x 2 martin datasci 4096 Jul  1 23:47 projects
# -rw-rw-r-- 1 martin datasci   11 Jun 26 10:15 report.txt
```

Let’s break down the first line (i.e., the `projects` folder):

+-------------------------------------------------------------------------------------+---------------------------------+----------------+-----------------+-------------------------------------------------+--------------------+----------------+
| File Mode                                                                           | Links                           | Owner          | Group           | Size                                            | Timestamp          | Name           |
+=====================================================================================+=================================+================+=================+=================================================+====================+================+
| **`drwxr-xr-x`**                                                                    | **`2`**                         | **`martin`**   | **`datasci`**   | **`4096`**                                      | **`Jul 1 23:47`**  | `projects`     |
|                                                                                     |                                 |                |                 |                                                 |                    |                |
| -   **`d`** entry is a **directory**                                                | 2 hard links:                   | **User** owner | **Group** owner | The space the folder occupies on the filesystem | Last modified date | Directory name |
|                                                                                     |                                 |                |                 |                                                 |                    |                |
| -   **`rwx`**: **r**ead, **w**rite, and e**x**ecute for **owner**                   | -   The directory itself (`.`)  |                |                 |                                                 |                    |                |
|                                                                                     | -   The parent reference (`..`) |                |                 |                                                 |                    |                |
| -   **`r-x`: r**ead + e**x**ecute for users in **group** *and* **others**; no write |                                 |                |                 |                                                 |                    |                |
+-------------------------------------------------------------------------------------+---------------------------------+----------------+-----------------+-------------------------------------------------+--------------------+----------------+

The second line displays the permissions for `report.txt`:

+-----------+---------+---------+---------+---------+-----------+---------+
| File Mode | Links   | Owner   | Group   | Size    | Timestamp | Name    |
+===========+=========+=========+=========+=========+===========+=========+
|           |         |         |         |         |           |         |
+-----------+---------+---------+---------+---------+-----------+---------+

1.  **`-`** – this entry is a **regular file**, not a directory or special.
2.  **`rw-` (owner)** – **r**ead and **w**rite allowed for **martin**; no execute.
3.  **`rw-` (group)** – **r**ead and **w**rite allowed for **datasci** group; no execute.
4.  **`r--` (others)** – **r**ead only for everyone else; no write or execute.
5.  **`1`** – a single hard link to this file.
6.  **`martin`** – user owner.
7.  **`datasci`** – group owner.
8.  **`11`** – file size is 11 bytes.
9.  **`Jun 26 10:15`** – last modified on June 26 at 10:15.
10. **`report.txt`** – the filename.

### Change Permissions

To adjust who can do what, use the <code>@sec-chmod</code> command. There are two common styles: **symbolic** (using letters) and **numeric** (using numbers). We’ll start with symbolic—it’s easier to read.

``` bash
ls -l data/README.md
# -rw-r--r--@ 1 username  staff  8834 May 14 09:33 data/README.md
```

The file permissions for `data/README.md` are stored in the combination of characters and symbols returned from `ls -l`:

Breakdown of **`-rw-r--r--@`**:

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -
```

The `-` indicates that the item is a regular file.
:::

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -rw-
```

The first three characters after the dash represent the user (owner) permissions. The user has read (`r`) and write (`w`) permissions but no execute (`x`) permission.
:::

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -rw-r--
```

The next three characters represent the group permissions. The group has read (`r`) permission only, with no write (`w`) or execute (`x`) permissions.
:::

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -rw-r--r--
```

The final three characters represent the permissions for others (everyone else). Like the group, others have only read (`r`) permission, with no write (`w`) or execute (`x`) permissions.
:::

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -rw-r--r--@
```

The `@` symbol indicates that the file has extended attributes, which are additional metadata stored by the operating system.
:::

::: {layout="[50,50]" layout-valign="top"}
``` {.sh style="background-color: #f8f8f8; font-style: italic;"}
# -rw-r--r--@ 1
```

The `1` indicates the count of hard links pointing to the file, which in this case is the file itself. This is typical for regular files that haven't been explicitly linked elsewhere.
:::

#### Symbolic Mode

This command adds (`+`) execute (`x`) permission to the user (`u`) of the file.

-   **u** = owner (you)
-   **g** = group
-   **o** = others
-   **a** = all (owner, group, and others)

And the operations:

-   **+** to add a permission
-   **-** to remove a permission
-   **=** to set exactly those permissions

**Examples:**

1.  **Add write permission for your group** on `report.txt`:

``` bash
chmod g+w report.txt
```

2.  **Remove execute permission for others** on `script.sh`:

``` bash
chmod o-x script.sh
```

3.  **Set read/write for owner, read-only for group and others** on `data.csv`:

``` bash
chmod u=rw,g=r,o=r data.csv
```

**Symbolic Notation**

To grant the permissions above (i.e., `-rw-r--r--`), we can group the arguments to `chmod`:

``` sh
chmod u=rw,go=r filename
```

In this case,

**`u=rw`**: specifies that the `u`ser (owner) can `r`ead and `w`rite

**`go=r`**: specifies that *both* the `g`roup and `o`thers can `r`ead, but not `w`rite or `e`xecute.

We can also use `chmod` to add permissions with `+`. The code below grants execute permissions to the user (owner) of the file named `data/README.md`:

``` sh
chmod u+x data/README.md
```

#### Numeric Mode

Each permission (r, w, x) has a numeric value:

-   **r** = 4
-   **w** = 2
-   **x** = 1

Add them together for each category:

| Permission | Value |
|------------|-------|
| none       | 0     |
| --x        | 1     |
| -w-        | 2     |
| -wx        | 3     |
| r--        | 4     |
| r-x        | 5     |
| rw-        | 6     |
| rwx        | 7     |

A three-digit number (e.g., **750**) sets owner, group, and others, in that order.

**Example:** To give the owner full access, the group read/execute, and others none:

``` bash
chmod 750 myscript.sh
```

-   Owner: 7 (rwx)
-   Group: 5 (r-x)
-   Others: 0 (---)

**Numeric Notation**

To set the permissions of the file `data/README.md` so that the user can read and execute it, the group can read it, and others have no permissions, we would use:

``` bash
chmod 750 data/README.md
```

Here, `7` stands for read (`4`), write (`2`), and execute (`1`) permissions for the user. `5` represents read and execute permissions for the group (`4+1`), and `0` means no permissions for others.

Both of these examples modify the permissions of `data/README.md` to enhance or restrict access as specified.

```{=html}
<!--
Overall, the permissions `644` (as represented numerically) allow the owner to read and write the file, while the group and others can only read it.
-->
```

### Permission Bits

Beyond basic read/write/execute, Linux has three **special bits** you may encounter:

1.  **Setuid (s on owner’s x)** When set on an executable file, this lets users run it with the file owner’s permissions.
2.  **Setgid (s on group’s x)** Similar to setuid, but for group permissions; on folders, new files inherit the folder’s group.
3.  **Sticky bit (t on others’ x)** On a folder, users can only delete or rename files they own, even if others have write permission.

You’ll see them as lowercase `s` or `t` in the permission string, for example:

``` plaintext
-rwsr-xr-x 1 root root 123K May 10  2023 /usr/bin/passwd
drwxrwxrwt 2 root root 4.0K Jun 20  2024 /tmp
```

-   `/usr/bin/passwd` uses **setuid** so you can change your password safely.
-   `/tmp` uses the **sticky bit** so everyone can write there, but you can only remove your own files.

------------------------------------------------------------------------

## A Simple Example

Let’s create a folder where members of the “team” group can freely share files, but nobody else can peek.

1.  **Create the group** (if it doesn’t exist):

``` bash
sudo groupadd team
```

2.  **Create the folder**:

``` bash
sudo mkdir /srv/team-files
```

3.  **Change the folder’s group** to `team`:

``` bash
sudo chown root:team /srv/team-files
```

4.  **Set permissions** so owner (root) and group can read/write/enter, others have no access:

``` bash
sudo chmod 770 /srv/team-files
```

5.  **Enable setgid** so new files inherit the group:

``` bash
sudo chmod g+s /srv/team-files
```

Now any member of the `team` group who drops files into `/srv/team-files` will automatically share them with the rest of the team.

## Check and Fix Permissions

**See current settings:**

``` bash
ls -ld /path/to/file-or-folder
```

**If you get “Permission denied”:**

1.  Check who owns the file and what group it’s in (`ls -l`).\
2.  Verify your user is in the right group (`groups`).\
3.  Adjust with `chmod`, `chown`, or `chgrp` as needed.

## Recap

Permissions are your first line of defense in Linux:

-   They protect your personal files.
-   They keep system files safe.
-   They let you share selectively.

With just a few simple commands—**`ls -l`**, **`chmod`**, **`chown`**, and **`chgrp`**—you’re in control of who can read, change, or run anything on your system. Practice by examining files on your machine, tweaking permissions in a safe folder, and watching how access changes. In the next chapter, we’ll dive into **working with processes**, showing you how Linux runs programs and how you can manage them.

`ls -l` output:

+-------------------+-------------------------------------------------------------------+
| Column            | Meaning                                                           |
+===================+===================================================================+
| **1. File mode**  | `d`/`-` + nine permission bits                                    |
+-------------------+-------------------------------------------------------------------+
| **2. Link count** | Number of hard links                                              |
+-------------------+-------------------------------------------------------------------+
| **3. Owner**      | User name who owns the file/directory                             |
+-------------------+-------------------------------------------------------------------+
| **4. Group**      | Group name associated with the file/directory                     |
+-------------------+-------------------------------------------------------------------+
| **5. Size**       | Size in bytes (for files) or filesystem-allocated size (for dirs) |
+-------------------+-------------------------------------------------------------------+
| **6. Timestamp**  | Last modification date and time                                   |
+-------------------+-------------------------------------------------------------------+
| **7. Name**       | File or directory name                                            |
+-------------------+-------------------------------------------------------------------+

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
