---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '#'
    
---

# Regular Expressions {#sec-regexp}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Regular expressions (or the singular 'regex') are powerful tools for searching and manipulating text data. A regex is made up of the symbols and characters that define specific patterns to be matched, identified or transformed. 

A regex operates on text--the sequence of characters that include letters, digits, punctuation, and other character types. Text serves as the 'data' or 'medium' for which the patterns the regex describes are searched.[^regex-data]

[^regex-data]: The `regex` patterns are not the data themselves, but rather a framework for locating or modifying text data. 

## Matching occurrences 

`+` matches one or more occurrences of the preceding element.

`[abc]`: Matches any one character listed (`a`, `b`, or `c`).

**Example**

Match any line containing one or more digits in `who_tb_data.csv.`:

```{bash}
grep -E '[0-9]+' data/raw/who_tb_data.csv
```

## Starts with

`^` matches the start of a line.

`[abc]`: Matches any one character listed (`a`, `b`, or `c`).

**Example**

Find lines beginning with an uppercase letter in `README.md`:

```{bash}
grep -E '^[A-Z]' data/README.md
```

## Ends with

`($| )`: followed by either end-of-line (`$`) *or* space


**Example**

Match lines that match a literal `ah` and either end-of-line (`$`) *or* space.

```{bash}
grep -E 'ah($| )' data/wu_tang.psv
```

## Search

**Example**

`^` matches the start of a line.

Find lines beginning with `Ghost` in `data/raw/wu_tang.csv`:

```{bash}
grep -E '^Ghost' data/raw/wu_tang.csv
```

**Example**

Print lines with `i` followed by one or more `l`s or lines with `i` followed by one or more `f`s:

```{bash}
grep -E  '(il+|if+)' data/wu_tang.psv
```

## Deep search

**Example**

Recursively search for `Kill` or `cybernetic` in any file under `data/raw`

```{bash}
grep -R -E '(Kill|cybernetic)' data/raw
```

**Example**

Print lines containing any four-digit sequence with square brackets and curly braces:

```{bash}
awk '/[0-9]{4}/ {print}' data/raw/who_tb_data.csv
```

Regular expressions can combine wildcards and special characters and are typically used with tools like <code>@sec-grep</code> (global regular expression print), <code>@sec-sed</code> (stream editor), <code>@sec-cat</code> (concatenate) and <code>@sec-awk</code>.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```

