---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# Set-Ups {#sec-setups .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```


```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

This section details the practical aspects of preparing your environment to work with a Linux system. We’ll cover options for setting up a virtual machine and working with shells and terminals. 

### Virtual Machines 

<strong>@sec-vms</strong> are software-based emulations of a physical computer that allow users to run Linux distributions in an isolated environment, utilizing virtualized hardware resources from the host system. 

```{=html}

<style>

.codeStyle span:not(.nodeLabel) {
  font-family: monospace;
  font-size: 1.75rem;
  font-weight: bold;
  color: #9753b8 !important;
  background-color: #faf9ed;
  padding: 0.2em;
}

</style>
```

```{mermaid}
%%| fig-width: 8.0
%%| fig-align: center
%%| fig-cap: 'Running Virtual Machines'
%%{init: {'theme': 'neutral', 'themeVariables': { 'fontFamily': 'monospace', "fontSize":"16px"}}}%%

flowchart TD
  VM("Virtual Machine")
  Linux("Linux Distro 🐧")
  HostSystem{"Host<br>Operating<br>System"}
  CPU("Virtual CPU")
  RAM("Virtual RAM 🧠")
  Disk("Virtual Disk 💽")

  HostSystem-->VM
  VM-->Linux
  VM-->CPU
  VM-->RAM
  VM-->Disk
  
    style HostSystem fill:#fefefe,font-size:16px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style VM fill:#,font-size:15px,stroke:#000,stroke-width:2px
    style Linux fill:#fefefe,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style CPU fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style RAM fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style Disk fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3

```

This chapter covers virtualization software options, various Linux distributions, and provides an example of setting up Ubuntu on VirtualBox. 

### Shells 

<strong>@sec-shells</strong> are the command-line interpreters that allow users to interact with the operating system by executing commands and running scripts.

```{=html}

<style>

.codeStyle span:not(.nodeLabel) {
  font-family: monospace;
  font-size: 1.75rem;
  font-weight: bold;
  color: #9753b8 !important;
  background-color: #faf9ed;
  padding: 0.2em;
}

</style>
```

```{mermaid}
%%| fig-width: 8.0
%%| fig-align: center
%%| fig-cap: 'Using Shells'
%%{init: {'theme': 'neutral', 'themeVariables': { 'fontFamily': 'monospace', "fontSize":"16px"}}}%%
flowchart TD
    Shell["Shell"]
    Shell --> Exec("Executes Commands")
    Shell --> Script("Runs Scripts")
    Exec --> OS["Host Operating System"]
    Script --> OS

    style Shell fill:#fefefe,font-size:14px,stroke:#000,stroke-width:2px,rx:3,ry:3
    style Exec fill:#fefefe,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style Script fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style OS fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    
```

### Terminals 

<strong>@sec-terminals</strong> are text-based interfaces that allow users to interact with the operating system by entering commands directly.

```{=html}

<style>

.codeStyle span:not(.nodeLabel) {
  font-family: monospace;
  font-size: 1.75rem;
  font-weight: bold;
  color: #9753b8 !important;
  background-color: #faf9ed;
  padding: 0.2em;
}

</style>
```

```{mermaid}
%%| fig-width: 8.0
%%| fig-align: center
%%| fig-cap: 'Using Terminals'
%%{init: {'theme': 'neutral', 'themeVariables': { 'fontFamily': 'monospace', "fontSize":"16px"}}}%%
flowchart TD
    User[User] --> Term["Terminal<br>Emulator"]
    Term --> Shell["Shell<br>(Command<br>Processor)"]
    Shell --> Prog["Other Programs"]
    Shell --> Out["Returns Output"]
    Prog --> Out


    style User fill:#,font-size:16px,stroke:#000,stroke-width:1px
    style Term fill:#fefefe,font-size:14px,stroke:#000,stroke-width:2px,rx:3,ry:3
    style Shell fill:#fefefe,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style Prog fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    style Out fill:#efefef,font-size:14px,stroke:#000,stroke-width:1px,rx:3,ry:3
    
```


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```

