---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '#'
    
---

# Special Characters {#sec-special-chars}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```



Some characters such as curly braces, single quotes, ampersand, and parentheses have special meanings in the shell. They can be used with arguments to add search, match, and print functionality.

## Braces: `{}` {#sec-braces}

**Brace Expansion** (`{}`) allows the creation of multiple text strings from a pattern containing braces. 

**Example**

List `.tsv` and `.csv` files named `wu_tang`:

```{bash}
ls data/wu_tang.{tsv,psv}
```

Expands into:

```{bash}
cat data/wu_tang.tsv 
```

```{bash}
cat data/wu_tang.psv
```

## Backslash: `\` {#sec-backslash}

`\` escapes the following character, nullifying its special meaning

**Example** 

Print text with spaces and the ampersand:

```{bash}
#| eval: false
echo "File name with spaces \& special characters"
# File name with spaces & special characters
```


## Single quotes: `''` {#sec-single-quotes}

Single quotes (`' '`) treat every character literally, ignoring the special meaning of all characters.

**Example** 

`echo '$HOME'` prints `$HOME`, not the path to the home directory:

```{bash}
echo '$HOME'
```

## Double quotes: `""` {#sec-double-quotes}

Double quotes (`" "`) allow for the inclusion of special characters in an argument, except for the dollar sign (`$`), backticks (\` \`), and backslash (`\`). 

**Example** 

`echo "$HOME"` prints the path to the home directory:

```{bash}
#| eval: false
echo "$HOME"
#> /Users/username
```

## Tilde: `~` {#sec-tilde}

`~` represents the home directory of the current user.

**Example** 

Print a folder `tree` of the items in the user's `Library/Python` directory:

```{bash}
#| eval: false
tree ~/Library/Python -L 1
#> /Users/username/Library/Python
#> └── 3.9
#>
#> 2 directories, 0 files
```


## Dollar Sign: `$` {#sec-dollar}

`$` indicates a variable.

**Example** 

`echo $PATH` prints the value of the `PATH` environment variable:

```{bash}
#| eval: false
echo $PATH
```


## Ampersand: `&` {#sec-ampersand}

`&` runs a command in the background.

**Example** 

`firefox &` opens Firefox in the background, allowing the terminal to be used for other commands.

```{bash}
#| eval: false
firefox &
```

## Semicolon: `;` {#sec-semicolon}

`;` separates multiple commands to be run in sequence.

**Example** 

Change the directory to `data/` and then print a folder `tree` of it's contents with the full path (`-f`), size (`-s`) in SI units (powers of 1000) in a more human readable way (`-h`).

```{bash}
#| eval: false
cd data/; tree -f -s --si -h
#> [ 320]  .
#> ├── [6.1k]  ./music_vids.tsv
#> ├── [ 12k]  ./pwrds.tsv
#> ├── [ 320]  ./raw
#> │   ├── [ 13k]  ./raw/ajperlis_epigrams.txt
#> │   ├── [6.5k]  ./raw/music_vids.csv
#> │   ├── [ 12k]  ./raw/pwrds.csv
#> │   ├── [1.3k]  ./raw/roxanne.txt
#> │   ├── [4.5k]  ./raw/trees.csv
#> │   ├── [4.8k]  ./raw/vg_hof.csv
#> │   ├── [ 381]  ./raw/who_tb_data.csv
#> │   └── [ 263]  ./raw/wu_tang.csv
#> ├── [9.0k]  ./README.md
#> ├── [4.4k]  ./trees.tsv
#> ├── [4.8k]  ./vg_hof.tsv
#> ├── [ 462]  ./wu_tang.psv
#> └── [ 263]  ./wu_tang.tsv
#> 
#> 2 directories, 15 files
```



## Greater Than: `>` {#sec-gt}

**Redirection operators**: `>` directs output to a file or a device. 

**Example** 

`echo "This is my 2nd file" > myfile2.txt` writes `"This is my 2nd file"` into `myfile2.txt`:

```{bash}
echo "This is my 2nd file" > myfile2.txt
```

## Less Than: `<` {#sec-lt}

**Redirection operators**: `<` takes input from a file or a device.

**Example** 

Then `wc < myfile2.txt` counts the words in `myfile2.txt`:

```{bash}
wc < myfile2.txt
```


## Parentheses: `()` {#sec-parens}

Parentheses can be used to group commands or for command substitution with `$( )`.

**Example** 

Runs `ls` in `/data` without changing the current directory:

```{bash}
(cd data; ls)
# confirm working directory hasn't changed 
pwd
```

`$(command)` uses the output of `command`.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
