---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: ' '
---

:::{style="color: #9b2e00; font-style: italic;"}

# Syntax {#sec-syntax .unnumbered}

:::

```{bash}
#| label: setup
#| include: false
touch myfile.txt
echo 'This is my file' > myfile.txt
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_construct
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under construction. Thank you for your patience."
)
```

In the previous section, we covered a few common commands used in Linux-like operating systems. In following chapters, we'll dive deeper into the syntax-- commands, arguments, and options--typed into the terminal or written in a shell script. 

We'll also discuss creating variables and how to chain commands together using pipes.

### Commands 

At the core of every Linux interaction is the command—an executable program or shell function that specifies the action the system should perform, such as listing directory contents with `ls` or launching complex workflows. The command is the first token that the shell parses, influencing how the rest of the line is interpreted. Whether located in `/bin`, a user-defined `$PATH`, or as a shell built-in, each command provides specific, reusable functionality that can be combined with arguments, options, redirections, and pipes to create effective one-liners or scripts.

### Arguments 

Arguments are the data items that follow a command and specify the objects it should operate on, such as files, directories, and IP addresses. The shell passes these arguments to the program as an ordered array (e.g., `$1`, `$2`, etc. in POSIX shell), allowing it to differentiate between multiple targets. Properly quoting arguments with whitespace or special characters is essential to avoid unexpected splitting or expansion, ensuring accuracy and security in both terminal work and scripts.

### Options 

Options, also known as "flags" or "switches," modify a command's default behavior without altering its core purpose. They are signified by a single dash for short forms (e.g., `-l`) or a double dash for longer forms (e.g., `--long`). Options can toggle modes, adjust verbosity, specify output formats, or enforce safety checks, and many can be combined in a single command (e.g., `ls -lah`). Familiarity with common option conventions—like `-h` for help, `-v` for verbose, and `-q` for quiet—streamlines troubleshooting and enhances command clarity.

### Pipes 

A pipe (`|`) connects the output of one command to the input of the next, creating a data stream that reflects the Unix philosophy of combining small, single-purpose tools. This allows efficient processing of large datasets, as seen with commands like `grep`, `cut`, and `sort` working together on log streams. Mastering pipes, along with redirection operators like `>`, `>>`, and `<`, facilitates quick analysis and automated processing workflows.


### Variables 

Variables act as named containers for data—like strings, numbers, or command substitutions—that scripts and interactive shells use to store information and provide context. Environment variables, such as `$PATH`, `$HOME`, and `$LANG`, influence system behavior, while local shell variables (declared with `VAR=value`) are session-specific unless exported (`export VAR`). Proper scoping, quoting (`"$VAR"`), and sanitation are crucial for secure scripting and align with best practices like the CIS Benchmarks for least privilege and configuration integrity.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
