---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '# '
---

# Universal interface {#sec-plain-text .unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Plain text files are simple, versatile, and easy to work with. They can be edited with any text editor and don't require specialized software, making actions transparent and learning accelerated. Plain text can also be easily manipulated using standard Linux text-processing tools such as **<code>@sec-grep</code>**, **<code>@sec-sed</code>**, and **<code>@sec-awk</code>**. With simple one-liners from the command line, users can search for a specific line, replace text across multiple files, or transform data formats.

## Text Editors and The Linux Philosophy

The Linux philosophy emphasizes simplicity, clarity, and the principle of “doing one thing well.” Plain text embodies this philosophy, serving as a simple, straightforward, and versatile means of interaction between the user, the system, and the programs running on it. This philosophy also underpins the design of Linux text editors, which range from the simple (like nano) to the powerful and extensible (like vi and emacs).

Text commands and editors are not just tools but the medium through which users communicate with the system and manipulate it to their will. Mastering these commands and editors opens up a world of possibilities for efficient system management, programming, and beyond.

This section will explore the core text commands that every Linux and Linux user should know, from file manipulation to text processing and searching. We will also introduce the most popular text editors, guiding you through their primary usage and highlighting their unique features. 


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
