---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
execute: 
  eval: false
---

# [Users and Groups]{style="color: #9b2e0c;"} {#sec-usr-grps}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
#| eval: true
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being reviewd Thank you for your patience.
  "
)
```

Linux is built for sharing. Whether you’re the only person on your laptop or part of a team on a shared server, Linux keeps everyone’s files and settings organized and secure by using **users** and **groups**. In this chapter, you’ll learn how Linux separates people and workloads, how to create and manage accounts, and how to use groups to let the right people share the right stuff.

## Why Users?

***Users in an operating system provide a transparent, secure and private personal space.***

Each user gets a "home" folder where their personal files live (i.e., `/home/john`). No one else can poke around in `home/john` without explicit permission. By default, one user can’t read or change another user’s files. This keeps us from accidentally breaking someone else’s work (and vice versa).

If something goes wrong (i.e., a file is deleted, a program crashes), we (Admins) can see which user account ran the command, which helps with troubleshooting.

## The Superuser

The **`root`** account is like the **master** user. `root` can do anything—install software, delete any file, change other users’ passwords. Because `root` has so much power, we usually do routine tasks as a normal user and only become `root` when we need more authority.

We can use **`sudo <COMMAND>`** ("superuser do") to run single commands as `root`, instead of logging in as the root `user` directly.

### Check Accounts

Open a terminal and try these:

```{bash}
whoami
# martin 
```

Lists your user ID (UID), your primary group ID (GID), and any extra groups you belong to.

```{bash}
id
# uid=1000 (martin) gid=1000(martin)groups=1000(martin),4(adm),24(cdrom),
# 27(sudo),30(dip),46(plugdev),100(users),114(lpadmin)
```

### List All Users

All user accounts are recorded in the file `/etc/passwd`. You can view it by running:

```{bash}
cut -d: -f1 /etc/passwd
```

```{verbatim}
root
daemon
bin
sys
sync
games
man
lp
mail
news
uucp
proxy
www-data
backup
list
irc
_apt
nobody
systemd-network
systemd-timesync
dhcpcd
messagebus
syslog
systend-resolve
uuidd
USbnux
tss
systemd-oom
kernoops
whoopsie
dnsmasq
avahi
tcpdump
sssd
speech-dispatcher
cups-pk-helper
fwupd-refresh
saned
geoclue
cups-browsed
hplip
polkitd
rtkit
colord
gnome-initial-setup
gdm
na-openvpn
gnone-remote-desktop
martin
```

That command prints the list of user names (the first field in each line).

> **Tip:** On a typical desktop you’ll see your account plus system accounts (like `daemon`, `syslog`). You usually only manage the human users.

### Add New Users

Log into the `su` account to manage user and groups to avoid typing `sudo` with the commands below. Linux provides a friendly command called **`adduser`** (on Debian/Ubuntu) or **`useradd`** (on Red Hat–based systems). We’ll use `adduser` here:

```{bash}
adduser bob
```

The terminal returns the following prompts:

1.  **Enter a password** for the new account.\
2.  **Confirm** the password.\
3.  Optionally, enter the user’s full name and other details (you can press Enter to skip).

```{verbatim}
New password:
Retype new password:
passwd: password updated successfully
Changing the user information for bob
Enter the new value, or press ENTER for the default
    Full Name []: Bob Smith
    Room Number []: 101
    Work Phone []: 555-555-5555
    Home Phone []: 555-555-5551
    Other []: 555-555-5552
Is the information correct? [Y/n] y
info: Adding new user 'bob' to supplemental / extra groups 'users' ...
info: Adding user 'bob' to group 'users' ...
```

When it’s done, you’ll have a new home folder at `/home/bob` and a private environment just for Bob.

### Change Passwords

If someone forgets their password, as admins we can reset it:

```{bash}
passwd bob
```

We’ll be asked to enter the new password for Bob and then confirm it.

### Remove Users

To delete an account and its home folder:

```{bash}
deluser --remove-home bob
```

**`--remove-home`** also erases `/home/bob`. If we only want to delete the account but keep the files, omit this option.

### Change user

As Admins, we can sign in as other users with `su` (*switch user*)

```{bash}
su bob
```

We can move into the user's home directory with `cd`: 

```{bash}
cd ~
```

Confirm this with `pwd`: 

```{bash}
pwd
```

```{verbatim}
/home/bob
```


## Groups

A **`group`** is a collection of users. Groups make it easy to share files:

**Example:** A 'design' group might include Alice, Bob, and Carol. Any files owned by the design group can be read or modified by all three.

Linux gives every user a **primary group** (by default, a `group` with the same name as their user). We can also add users to extra (secondary) groups.

We'll switch back to our original user to see what groups we belong to: 

```{bash}
bob@martin-VirtualBox:-$ exit
exit # out of bob into root
root@martin-VirtualBox:/home/martin# exit  
exit # out of root into original user
martingnartin-VirtualBox:-$
```

### View Groups

To see which groups we’re in:

```{bash}
groups
```

```{verbatim}
martin adm cdrom sudo dip plugdev users lpadmin
```

To see which groups a user is in:

```{bash}
groups bob
```

```{verbatim}
bob : bob users
```

As we can see, `bob` is a user who belong to the `users` group (also named `bob`).

To list all groups on the system:

```{bash}
cut -d: -f1 /etc/group
```

```{verbatim}
root 
daemon 
bin 
sys
adm 
tty 
disk
lp
mail 
news
uucp
man 
proxy 
kmem
dialout 
fax 
voice 
cdron 
floppy 
tape 
sudo 
audio 
dip 
www-data 
backup 
operator 
list 
irc
src
shadow 
utmp
video 
sasl 
plugdev
staff 
games 
users
nogroup 
systemd-journal 
systemd-network
crontab 
systemd-timesync 
input 
sgx
kvn
render 
messagebus 
syslog
systemd-resolve
_ssh
tss
ssl-cert 
systemd-oom
bluetooth 
rdma
whoopsie 
netdev
avahi 
tcpdump
sssd
lpadmin
fwupd-refresh
scanner 
saned 
geoclue 
pipewire 
polkitd
rtkit
colord 
gdr
nm-openvpn
lxd
gnome-remote-desktop 
gamemode 
gnome-initial-setup 
martin 
bob
```

### Create New Group

Sign into `su` to add groups:

```{bash}
sudo su
```

```{bash}
groupadd datasci
```

Creates a group called `datasci`.

### Add Users to a Group

Add `bob` to the `datasci` group:

```{bash}
usermod -aG datasci bob
```

The `-aG` flag means "*append to the supplemental groups.*"

To confirm, run:

```{bash}
groups bob
```

```{verbatim}
bob : bob users datasci
```

We should see `datasci` listed.

### Remove User from a Group

Sometimes we need to take someone off a team:

```{bash}
gpasswd -d bob datasci
```

Removes Bob from the `datasci` group.

## Ownership and Permissions

This is a brief overview of ownership and permissions, but we'll dive deeper in the [Permissions](permissions.qmd) chapter.

Files in Linux has three categories of permissions:

| Permissions | Description                |
|-------------|----------------------------|
| **Owner**   | the user who owns the file |
| **Group**   | the file's group           |
| **Others**  | everyone else              |


 And three permission types:

| Permission Type | Description                               |
|-----------------|-------------------------------------------|
| `r`             | read                                      |
| `w`             | write                                     |
| `x`             | execute (run a program or enter a folder) |


We'll sign in as `bob` and create a new `report.txt` file in `/home/bob/`: 

```{bash}
su bob
cd ~
echo "new report" >> report.txt
```

We can see permissions with:

```{bash}
ls -l report.txt
```

For example:

```{verbatim}
-rw-rw-r-- 1 bob bob 11 Jun 26 10:15 report.txt
```

-   `-rw-rw-r--` means:

    -   Owner (bob) can read/write.
    -   Group (bob) can read/write.
    -   Others have read access.

-   `bob` is the owner and the group.

### Change Ownership

**`chown`** changes the file’s owner:

```{bash}
chown martin report.txt
```

### Change Group

**`chgrp`** changes the file’s group:

```{bash}
chgrp datasci report.txt
```

### Change Both

Change both at once with **`chown`**:

```{bash}
chown martin:datasci report.txt
```

### Change Permissions

To adjust who can do what:

```{bash}
chmod g+w report.txt
```

Adds write permission (`+w`) to the group (`g`).

```{bash}
chmod o-rwx report.txt
```

Removes all rights (`-rwx`) from others (`o`).

> **Tip:** We can also use numeric codes (e.g., `chmod 660 report.txt` gives read/write to owner and group, none to others), but starting with letters (u, g, o) is more intuitive.

## A Simple Example

We'll set up a shared folder for the "`datasci`" group:

**Create the folder:**

```{bash}
mkdir /srv/datasci-projects
```

**Change its group and owner:**

```{bash}
chown root:datasci /srv/datasci-projects
```

**Set group write permissions:**

```{bash}
chmod 770 /srv/datasci-projects
```

**Enable the "setgid" bit** so new files inherit the group:

```{bash}
chmod g+s /srv/datasci-projects
```

Now any member of the `datasci` group can drop files in `/srv/datasci-projects`, and they’ll automatically belong to the `datasci` group.

## Troubleshoot

-   **"Permission denied"** User doesn't have the right permission. Check `ls -l` and see if they need to be in a group, change ownership, or use `sudo`.

-   **User can’t access shared folder** Make sure:

    1.  The folder’s group matches the user’s group.
    2.  Permissions allow group read/write (`rwx` or at least `rw-`).
    3.  The user has been added (and re-logged-in) to that group.

-   **"User already exists"** We might have a system or service account with that name. Pick a unique user name.

## Recap

Users and groups are Linux’s way of keeping things tidy and safe. By:

-   Creating individual accounts,
-   Assigning people to groups,
-   Setting permissions on files and folders,

As Admins, we control who can see, change, or run anything on the system. Once we’re comfortable with these basics, we can:

-   Set up project teams,
-   Secure sensitive data,
-   Delegate system tasks safely.

In the next chapter, we’ll explore permissions in depth. 
