---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '#'
    
---

# Wildcards {#sec-wildcards}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Wildcards (or *glob patterns*) can be used with commands to match filenames, paths, or filter text (`ls`, `cp`, `mv`, `rm`, etc.). The shell expands arguments and wildcards into a list of files or directories that match the pattern.

## Asterisk: `*` {#sec-wc-asterisk}

`*` is a wildcard for matching **zero or more characters**. 

**Example** 

List all files in the `data/` directory that end with `.md`:

```{bash}
ls data/*.md
```

**Example** 

List all top-level files starting with `w` and ending in `.tsv` in `data/`

```{bash}
ls data/w*.tsv
```

**Example**

Print the top ten lines of the `pwrds` .csv file in `data/raw/` via `*`:

```{bash}
head data/raw/p*.csv
```


## Question Mark: `?` {#sec-wc-question}

`?` is the wildcard for matching **exactly one character**.

**Example** 

`ls myfile?.txt` lists files like `myfile2.txt`, but not `myfile.txt` and `my file 3.txt`:

```{bash}
ls myfile?.txt
```

**Example** 

List five-character names in` data/raw/` ending with `.csv`:

```{bash}
ls data/raw/?????.csv
```

**Example** 

List any four-character names in` data/raw/` ending with `s.csv`:

```{bash}
ls data/raw/????s.csv
```


## Square brackets: `[]` {#sec-wc-square-brackets}

`[abc]`: Matches any one character listed (a, b, or c).

**Example** 

`[a-z]`: match the top-level TSV files starting with `m`, `p`, or `t` in `data/`:

```{bash}
ls data/[mpt]*.tsv
```

**Example** 

Matches files beginning with `p` or `t` and ending in `.tsv` in `data/`:

```{bash}
ls data/[pt]*.tsv
```


**Example** 

Matches `ajperlis_epigrams.txt` (but excludes` roxanne.txt`) in `data/raw/`:

```{bash}
ls data/raw/[!r]*.txt
```


**Example** 

Match any one character in range (`a` to `p`).

```{bash}
ls data/[a-p]*
```

In the next section we'll cover how to use special characters to enhance matching and searching text with tools like `grep`, `cat`, `head`, `awk`. etc.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```