---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# Windows {#sec-wsl}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being reviewed. Thank you for your patience."
)
```

Windows Subsystem for Linux ([WSL](https://learn.microsoft.com/en-us/windows/wsl/install)) allows you to run Linux on your Windows computer without a virtual machine or dual-booting. You can use Linux command-line tools and applications alongside your Windows programs.

## Install

Open Windows PowerShell as Administrator and run the following commands: 

```powershell
wsl install 
```

To view a list of available linux distributions (*distro*), run the following:

```powershell
wsl --list --online
```

Pick the distro you'd like to install (i.e., Ubuntu, Debian, Fedora, openSUSE, etc.) and install it using (replacing `<distro>` with the name of the distribution): 

```powershell
wsl --install -d <distro>
```

See example below: 

```powershell
wsl --install -d Ubuntu
```

## Configure 

A Linux terminal looks like a text window with a prompt, something like:

```bash
john@LAPTOP:~$
```

* **`john`** is your Linux user name.
* **`LAPTOP`** is your computer’s name.
* **`~`** means you’re in your “home” folder (your personal space).
* After the dollar sign (`$`), you type commands.

###  Keeping Linux Updated

Linux uses a package manager to install and update software. 

**On Ubuntu or Debian**:

```bash
sudo apt update
```

This asks Windows for your Linux password (`sudo` means *run as administrator*).

After updating `apt`, download and install any available updates:

```bash
sudo apt upgrade
```

Run these commands every few weeks to keep things fresh and secure.

### Sharing Files 

One of WSL’s coolest features is file sharing:

**Windows → Linux**

Your Windows drives are mounted under `/mnt`. For example, your C: drive is at `/mnt/c`. So if you want to work on a file stored in Windows’ Documents folder, you might type:

```bash
cd /mnt/c/Users/YourName/Documents
ls
```

**Linux → Windows**

You can save files in your Linux home (`~/`), then open them in Windows by navigating in File Explorer to:

```powershell
\\wsl$\Ubuntu\home\john
```

(Replace `Ubuntu` and `john` with your distribution name and user name.)

### Customizing Shell

If you're spending a lot of time in the terminal, it's nice to customize it to your liking. Below are some examples:

**Change the prompt color** 

Edit a file called `.bashrc`. In your terminal, type:

```bash
nano ~/.bashrc
```

Scroll to the bottom and add a line like:

```bash
PS1="\[\e[32m\]\u@\h:\w\$ \[\e[0m\]"
```

Press Ctrl+O to save, Enter to confirm, then Ctrl+X to exit. Finally, reload with:

```bash
source ~/.bashrc
```

**Auto-update on launch** 

In the same file (`~/.bashrc`), you could add:

```bash
sudo apt update && sudo apt upgrade -y
```

Now every time you open your terminal, it checks for updates.

## Recap

Linux can feel a bit mysterious if we're used to clicking icons in Windows. WSL lets you run a "real" Linux command line inside Windows, without setting up a separate computer or virtual machine. WSL allows us to use Linux tools from the same computer we're already using.

WSL is fast, lightweight, and it integrates quickly with Windows files and applications.