---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# 🏗️ [Windows]{style="color: #9b2e0c;"} {#sec-wsl}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under contruction. Thank you for your patience."
)
```

Windows Subsystem for Linux ([WSL](https://learn.microsoft.com/en-us/windows/wsl/install)) allows you to run Linux on your Windows computer without a virtual machine or dual-booting. You can use Linux command-line tools and applications alongside your Windows programs.

## Install

Open Windows PowerShell as Administrator and run the following commands: 

```powershell
wsl install 
```

To view a list of available linux distributions (*distro*), run the following:

```powershell
wsl --list --online
```

Pick the distro you'd like to install (i.e., Ubuntu, Debian, Fedora, openSUSE, etc.) and install it using (replacing `<distro>` with the name of the distribution): 

```powershell
wsl --install -d <distro>
```

See example below: 

```powershell
wsl --install -d Ubuntu
```

## Configure 



## Recap